Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-03-27 demystify.tex
updates

Second, PhD students in CS \emph{do not} need to worry about funding, especially at good R1
universities in the US. If you are admitted, you will almost certainly
receive \emph{full funding} to support your study, including tuition,
health insurance, and stipend (i.e., you get paid for your study). Moreover, depending on the university,
you may even receive additional benefits such as summer pay, laptops, (conference/workshop) traveling. \autoref{sec:funding} provides more details on funding.

@@

\paragraph{Why CS depts do not waive application fee?}  Note that some universities do waive, e.g., for domestic students (\autoref{sec:domestic-students}), however most do not.  In any case, application fee is typically a requirement of the university. Individual departments and programs do not have the flexibility to waive the application fee, even if they want to.

In my opinion, requiring applicants to pay the fee helps ensure their seriousness, as it filters out non-serious candidates. Also, if the application process were free for everyone, most good CS programs would receive an overwhelming number of applications to review.

@@

\chapter{Research Opportunities}\label{sec:research-opportunities}

@@

 This section provides some guidance on how to gain research experience, e.g., as an undergraduate student.  It also give advice to students from small colleges where research opportunities are limited.

@@

There might be also special programs for undergraduate research from the university, e.g., GMU has the OSCAR program, UNL has UCARE, NSF has REUs (Research Experience for Undergraduates) for US citizens and permanent residents.

@@

\chapter{Cultural Differences and Other Issues}\label{sec:cultural}

This section lists some cultural issues that international students might want to pay attention to. These issues apply in general and not just PhD admission.

@@

Students, especially international ones, often get confused about what to do after they get an offer from a professor, e.g., is it OK to postpone or accept/reject offers?, and most importantly, how to do so without offending anyone. Many also wonder about accepting an offer and then later rejecting it.

>>>>

Second, PhD students in CS \emph{do not} need to worry about funding, especially at good R1
universities in the US. If you are admitted, you will almost certainly
receive \emph{full funding} (\autoref{sec:funding}) to support your study, including tuition,
health insurance, and stipend (i.e., you get paid for your study). Moreover, depending on the university,
you may even receive additional benefits such as summer pay, laptops, (conference/workshop) traveling.

\autoref{sec:summary} provides an overview of the application process, which you can explore in further details in the following chapters.

@@

\paragraph{Why CS depts do not waive application fee?}  Some universities do waive, e.g., Rice for all PhD applicants and many universities for domestic students (\autoref{sec:domestic-students}), however most do not.  In any case, application fee is typically a requirement of the university. Individual departments and programs do not have the flexibility to waive the application fee, even if they want to.

In my opinion, requiring applicants to pay the fee helps ensure their seriousness, as it filters out non-serious candidates.  Also, most CS programs already receive many applications, and would be overwhelmed if the application process were free.

@@

\chapter{Research Opportunities}\label{sec:research-opportunities}
\begin{center}
    \includegraphics[scale=0.5]{files/phd100404s.png}
  \end{center}

@@

 This section provides some guidance on how to gain research experience, e.g., as an undergraduate student.  It also gives advice to students from small colleges where research opportunities are limited.

@@

There might be also special programs for undergraduate research from the university, e.g., GMU has the OSCAR program, UNL has UCARE, NSF has REUs (for US citizens and permanent residents).

@@

\chapter{Cultural Differences}\label{sec:cultural}

This section lists some cultural issues that students, especially international ones, might want to pay attention to. These issues apply in general and not just PhD admission.

@@

Students often ask about what to do after they get an offer from a professor, e.g., is it OK to postpone or accept/reject offers?, and most importantly, how to do so without offending anyone. Many also wonder about accepting an offer and then later rejecting it.
